---
title: "Prediction Assignment Writeup 2"
author: "Jackie Orr"
date: "April 14, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Overview
This analysis will observe 6 participants performing the activity of lifting dumbells correctly and incorrectly while wearing exercise tracking devices. We will examine how well the activity trackers assess how well the participants perform the exercise.

##Install Packages


```{r, echo=TRUE}
library(lattice);
library(caret);
library(rattle);
library(rpart.plot);
library(ggplot2);
library(randomForest)
```

##Load Data


```{r, echo=TRUE}
url.train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url.test <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
wholedata <- read.csv(url(url.train), na.strings = c("NA", "", "#DIV0!"))
testingSet <- read.csv(url(url.test), na.strings = c("NA", "", "#DIV0!"))
```

##Cleaning Data
```{r, echo=TRUE}
NA_Count = sapply(1:dim(wholedata)[2],function(x)sum(is.na(wholedata[,x])))
NA_list = which(NA_Count>0)
wholedata = wholedata[,-NA_list]
wholedata = wholedata[,-c(1:7)]
wholedata$classe = factor(wholedata$classe)

NA_Count1 = sapply(1:dim(testingSet)[2],function(x)sum(is.na(testingSet[,x])))
NA_list1 = which(NA_Count1>0)
testingSet = testingSet[,-NA_list]
testingSet = testingSet[,-c(1:7)]
dim(wholedata)
```

```{r, echo=TRUE}
  dim(testingSet)
```

Test set will be used for final model validation
```{r, echo=TRUE}
    inTrain=createDataPartition(y=wholedata$classe, p=0.6, list=FALSE)
    training <-wholedata[inTrain,]
    testing <- wholedata[-inTrain,]
```

##Tree Method Model
```{r, echo=TRUE}
modfit1 <- train(classe ~.,method='rpart',data=training)
    fancyRpartPlot(modfit1$finalModel) 
```

```{r, echo=TRUE}
 pred=predict(modfit1,newdata=testing)
    z=confusionMatrix(pred,testing$classe)
    z$table
```
```{r, echo=TRUE}
 z$overall[1]
```

Model accuracy is 0.49, which indicates there is no purity in the model.

##Random Forest Model
```{r, echo=TRUE}
modfit2=randomForest(classe~., data=training, method='class')
    pred2 = predict(modfit2,testing,type='class') 
    qplot(roll_belt, magnet_dumbbell_y, colour=classe, data=training)
```

This model provides 99% accuracy and we will use it for predictions.

```{r, echo=TRUE}
   z2=confusionMatrix(pred2,testing$classe)
    z2$table
```
```{r, echo=TRUE}
    z2$overall[1]
```

##Conclusion
The Random Forest Method provides the highest accuracy and is considered the best fit for predicting this data set.
```{r, echo=TRUE}
   pred3 =  predict(modfit2,testingSet,type='class')
    nofiles = length(pred3)
    for (i in 1:nofiles){
        filename =  paste0("problem_id",i,".txt")
        write.table(pred3[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
    pred3
```

